---
title: "Assignment3"
output: html_document
date: "2023-10-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Using SQL in R

### Assignment Questions

##### For this assignment, I will answer the following two questions using charts and plots:

1.  Which geographic regions generate the most sales?
2.  What are the most profitable products?

Below are the steps. 

### Installing packages

`library(tidyverse)`

`library(RODBC)`

```{r include=FALSE}
library(sqldf)
library(tidyverse)
library(RODBC)
```

### Establishing Connection to SQL database: AdventureWorks2022

```{r}
db_conn <- odbcConnect("myDatabase", rows_at_time =1)
```

### Defining the SQL Query

To calculate the profit of a product, I assumed that Profit = SalesRevenue - AcquisitionCost

SalesRevenue is assumed to be UnitPrice in Sales.SalesOrderDetail. 

AcquisitionCost is assumed to be the StandardCost in Production.Product. However, some products seem to be purchased from external producers. Their purchasing price can be different from their StandardCost. I used StandardCost for every product in my calculation. 

Sales are aggravated from 2011 to 2014. 

```{r}
queryProductionTable <- "
-- Use AdventureWorks2022
Select sd.ProductID
	,pp.[Name] as ProductName
	,LineTotal as SalesAmt
	-- Revenue: group_by(ProductID), sumarize LineTotal
	,pp.StandardCost as ProductCost  
	-- Cost: summarize StandardCost by Mean, not Sum
	,Territory
	,OrderDate as SalesYear
	,TotalSalesQty as SalesQty
	,TotalStocked as ProductStocked
	,TotalOrdered as ProductOrdered
from
(SELECT ProductID
	,[Group] as Territory
	,YEAR(soh.OrderDate) as OrderDate
	,SUM(sod.OrderQty) as TotalSalesQty
	,SUM(sod.LineTotal) as LineTotal
  FROM (
	  Sales.SalesOrderDetail as sod
	  left join 
	  Sales.SalesOrderHeader as soh
	  on sod.SalesOrderID = soh.SalesOrderID
	  left join
	  Sales.SalesTerritory as st
	  on st.TerritoryID = soh.TerritoryID
	  )
  group by ProductID, OrderDate, [Group]
  ) sd 

 left join

  (SELECT ProductID
	,sum(stockedqty) as TotalStocked
	,sum(orderqty) as TotalOrdered
	,sum(ScrappedQty) as TotalScrapped
  FROM Production.WorkOrder
  group by ProductID
  -- order by ProductID
  ) p
  on sd.ProductID = p.ProductID
  
  left join

  Production.Product as pp
  on pp.ProductID = sd.ProductID
"
```

### Executing the SQL Query

```{r}
productionTable <- sqlQuery(db_conn, queryProductionTable,
                            stringsAsFactors=FALSE)
```

### Plotting in Rstudio

```{r}
productionTable |> 
  group_by(Territory) |> 
  summarize(totalSalePerc = sum(SalesAmt)/sum(productionTable$SalesAmt)) |> 
  # Rerrange Territories in a particular order so that Pacific comes first.
  mutate(Territory = 
           factor(Territory, 
                  levels = c("Pacific", "Europe", "North America"))) |> 
  
  # Preparing perc labels for pie chart. Input: 0.722 Output: 72%. 
  mutate(totalSalePercLabel = scales:: percent(round(totalSalePerc,2))) |> 
  
  ggplot(aes(x="", y=totalSalePerc, fill = Territory)) +
  geom_col() + 
  # pie chart command
  coord_polar(theta = "y") +
  
  # colouring the pie chart
  scale_fill_manual(values = c("grey","orange","navy")) +
  
  labs(title = "TOTAL REGIONAL SALES") +
  geom_text(aes(label = totalSalePercLabel),
            color = "white",
            size = 5,
            # must have for pie chart
            position = position_stack(vjust=0.5)) + 
  geom_text(aes(label = Territory
                # adjust position of in-chart labels
                ,vjust=-1
                ,hjust=0.6),
            color = "white",
            size = 5,
            position = position_stack(vjust=0.5)) +
  
  # Cleaning up the backgrounds
  theme(axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        title = element_text(size = 15), 
        
        # centering the title
        plot.title = element_text(hjust=0.5),
        
        # Remove legend
        legend.position = "none",
        
        # legend.text = element_text(size=13),
        panel.background = element_rect(fill = "white"))

testFill <- c("TotalProfit", "Totalsales")

productionTable |> 
  group_by(ProductName) |> 
  summarize(TotalSales = sum(SalesAmt),
            TotalProfit = 
              sum(SalesAmt) - (mean(ProductCost) * sum(SalesQty)))|> 
  arrange(desc(TotalProfit))|> 
  head(10) |> 
  ggplot(aes(x=ProductName, y= TotalProfit, TotalSales)) +
  geom_col(fill="navy", width = 0.5) +
  coord_flip() +
  
  # display large numbers properly.
  scale_y_continuous(labels=scales::comma) +
  
  labs(title="Most Profitable Products", y = "Total Profit", x="") +
  theme(axis.title.x = element_text(size=13),
        title = element_text(size = 15), 
        
        # centering the title
        plot.title = element_text(hjust=0.5),
        
        panel.background = element_rect(fill = "white"))
```

### Closing the connection to SQL Server
```{r}
odbcClose(db_conn)
```
